---
title: "DV_index_cleaning"
author: "Cande Torres"
date: "11/26/2021"
output: html_document
---

```{r}
library(tidyverse)
library(tidyr)
library(readr)
library(ggplot2)
library(dplyr)
library(broom)
library(readxl)
library(readr)
library(janitor)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Upload data 

```{r}
covid_OX<-read_csv("https://raw.githubusercontent.com/OxCGRT/covid-policy-tracker/162b5cae62b14993c1f0eda850aae5ee6001fdcb/data/OxCGRT_withnotes_2020.csv")
```

```{r}
#We actually don't need this data set because the Oxford tracker has that data. 
covid_impact<- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/owid-covid-data.csv")
```

```{r}
library(readxl)
covid_school <- read_excel("~/Desktop/Fall21/Advanced Data Science/KA_CT_project/Multivariate_data_2.0.xlsx",col_types = c("text", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric", "numeric", "numeric"))
```

# Clean

```{r}
covid_OX<- clean_names(covid_OX)
```

```{r}
covid_school<- clean_names(covid_school)
```

```{r}
covid_school<-covid_school %>% 
  select(-x16,-x17,-x18,-x19,-x20,-x21,-x22,-x23,-x24,-x25,-x26,-x27)
```

```{r}
covid_OX<- covid_OX %>% 
  select("country_name", "date", "c1_school_closing", "c2_workplace_closing", "c3_cancel_public_events", "c4_restrictions_on_gatherings", "c5_close_public_transport", "c6_stay_at_home_requirements", "c7_restrictions_on_internal_movement", "c8_international_travel_controls", "e1_income_support", "e2_debt_contract_relief", "e3_fiscal_measures", "e4_international_support", "h1_public_information_campaigns", "h2_testing_policy", "h3_contact_tracing", "h4_emergency_investment_in_healthcare", "h5_investment_in_vaccines","h6_facial_coverings","h7_vaccination_policy", "h8_protection_of_elderly_people", "confirmed_cases", "confirmed_deaths", "stringency_index", "government_response_index", "economic_support_index")
```

# Join data sets 

```{r}
data_join<-left_join(covid_school,covid_OX, by=c("unicef_country"="country_name"))
```

# DV: Level of Prioritization of education 

Note: avoid collinearity. Create index adding up variables that respond to country's measures attempting to make education a priority on their 2020 agenda. Because it's a time series dataset, choose a specific date to select, justify selection. Maybe 3 dates to show variation. 

- I will filter the data for the 15th of each month after July 2020 until December 2020.

```{r}
data_date<- data_join %>% 
  filter(date %in% c(20200715,20200815,20200915,20201015,20201115,20201215))
```

- Rank the UNICEF vars 
```{r}
#Code to create ranking for the UNICEF vars. 
UNICEF_ranked<-data_date %>% 
  mutate(across(c(`days_fully_open`,`instruction_days`), #for these vars, the higher the score, the higher the rank, but that is not the case for all vars 
    ~min_rank(.x))) 
```

```{r}
#For days_fully_closed, days_academic_break and days_partially_closed, the higher the score, the lower the rank. 
UNICEF_ranked<- UNICEF_ranked %>% 
  mutate(across(c(`days_fully_closed`, `days_academic_break`, `days_partially_closed`),
         ~min_rank(-.x)))
```

#raname long vars

```{r}
names(UNICEF_ranked)[names(UNICEF_ranked) == "support_to_girls_and_vulnerable_groups_children_with_disabilities_children_on_the_move_etc"] <- "Support_vulnerable"
names(UNICEF_ranked)[names(UNICEF_ranked) == "teacher_support_pedagogical_psychosocial_financial_etc"] <- "teacher_support"
names(UNICEF_ranked)[names(UNICEF_ranked) == "coverage_of_curriculum_subjects_and_content_planned"] <- "curriculum_coverage"
names(UNICEF_ranked)[names(UNICEF_ranked) == "remote_learning_content_sensitive_to_needs_of_different_groups"] <- "content_need_sensitive"
names(UNICEF_ranked)[names(UNICEF_ranked) == "quality_of_content_in_remote_learning_platforms"] <- "quality_remote_platform_content"
```

#Weighting index 

Recode UNESCO +C1 var to match UNICEF scale

```{r}
data_scaled<-UNICEF_ranked %>% 
  mutate(across(c(`tracking_or_monitoring_of_student_learning`, `quality_remote_platform_content`, `Support_vulnerable`, `content_need_sensitive`, `parental_caregiver_support`, `curriculum_coverage`, `teacher_support`, `access_to_remote_learning`, `quality_remote_platform_content`, `c1_school_closing`),
      ~.x*1000))
```

#re-scaling the other vars just in case 

```{r}
data_scaled<-data_scaled %>% 
  mutate(across(c(`c2_workplace_closing`, `c3_cancel_public_events`, `c4_restrictions_on_gatherings`, `c5_close_public_transport`, `c6_stay_at_home_requirements`, `c7_restrictions_on_internal_movement`, `c8_international_travel_controls`),
      ~.x*1000))
```
```{r}
data_scaled<-data_scaled %>% 
  mutate(across(starts_with("h"),
      ~.x*1000))
```
```{r}
data_scaled<-data_scaled %>% 
  mutate(across(c(`e1_income_support`, `e2_debt_contract_relief`, `e3_fiscal_measures`, `e4_international_support`),
      ~.x*1000))
```


#Coding Index 

```{r}
data_scaled<- data_scaled %>% 
  mutate(DV_Edu_prior_index= days_fully_open+
           instruction_days+
           days_fully_closed+
           days_academic_break+
           days_partially_closed+
           tracking_or_monitoring_of_student_learning+
           quality_remote_platform_content+
           Support_vulnerable+
           content_need_sensitive+
           parental_caregiver_support+
           curriculum_coverage+
           teacher_support+
           access_to_remote_learning+
           quality_remote_platform_content+
           c1_school_closing)
```

# Drop NAS
```{r}
data_clean<- data_scaled %>% 
  filter(!is.na(DV_Edu_prior_index))
```

# Add Cathegoric Leveling for Index

```{r}
summary(data_clean$DV_Edu_prior_index)
```

```{r}
data_DV_clean <- data_clean %>%
  mutate(category=cut(DV_Edu_prior_index, breaks=c(-Inf, 22000, 25400, Inf), labels=c("low","middle","high")))
```

#Write new data set 

```{r}
library(writexl)
```

```{r}
write.csv(data_DV_clean,"ADS_project_data_DV.csv")
write_xlsx(data_DV_clean,"ADS_project_data_DV.xlsx")
```


